---
title: "Mock Swiftkey Prediction Capstone"
author: "MT"
date: "2/26/2021"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE, cache=T}
knitr::opts_chunk$set(echo = FALSE, comment=NA)
```

## Introduction

Data Science Capstone project requires us to fulfi few components of the porject.

 - Exploring Text Datasets
 - Explore and create a model for nature language processing
 - Creating a Shiny app to display the output of our predictor
 - Creating this presentation to introduce the project and how the  app works.


## Exploration of Data

- Datasets were obtain from news, twitter  and blog posts
- Due to limitations for this project, only  1% of the data is randomly selected
- The 1% is than  processed and splited into training, test, and holding set. 
- Models were created from the training set, validated  to function on the holding set, before generating the  test results. 

## Data Cleaning

- Each of the data sets were cleaned and remove words or symbols that may not helpful to generate the ngrams
- Quanteda packaged was used along with other packages to tidy up the data
- Ngrams of two, three, four and five were generated in this project

## Results of the test set
- According to one of the resources perplexity is used to assess the effectiveness of the model. With lower values to be better.


```{r cars, echo = F}
library(data.table)
x<- data.table(model=c("Bigram", "Trigram", "Fourgram", "Fivegram"), perplexity=c("66.09", "8.87", "1.05", "0.28"))
x
```

## Limitations
- Only single model is applied here due to time constraint
- Other NLP models should be explored to validate the accuracy of current model used.

