---
title: "12_tidy_data"
author: "MT"
date: "7/12/2021"
output: html_document
---

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
table1
table2
table3
table4a
```

```{r}
table1 %>%
        mutate(rate=cases/population*10000)

table1 %>%
        count(year, wt=cases)

ggplot(table1, aes(year, cases)) +
        geom_line(aes(group = country), colour = "grey50") +
        geom_point(aes(colour = country))

```
###12.2.1 Exercises

####Question 1
table1. Observations of year, cases and population variables for each country. With multiple years in the same variable
table2. Observations of year, type of data and count as variabl for each country. Each data is sorted by country followed by data type
table3. Observation of rates group by country followed by their year observed. 
table4a. Observation of counts with year as the variable and sorted by coutnry.

####Question 2 and 3
```{r}
table2
plot2<- table2 %>%
        pivot_wider(names_from= type, values_from= count) %>%
        mutate(rate=cases/population*10000)
plot2
ggplot(plot2, aes(year, cases)) +
        geom_line(aes(group = country), colour = "grey50") +
        geom_point(aes(colour = country))
        
```

###Seperate
```{r}
table3 %>%
        separate(rate,into=c("cases", "populaiton"), convert=T)
```

###Unite
```{r}
table5
table5 %>%
        unite(new,century, year, sep="")

```

###Missing Values
```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr    = c(   1,    2,    3,    4,    2,    3,    4),
  return = c(1.88, 0.59, 0.35,   NA, 0.92, 0.17, 2.66)
)

stocks %>%
        pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(
    cols = c(`2015`, `2016`), 
    names_to = "year", 
    values_to = "return", 
    values_drop_na = TRUE
  )

stocks %>% 
  complete(year, qtr)

```

```{r}
treatment <- tribble(
  ~ person,           ~ treatment, ~response,
  "Derrick Whitmore", 1,           7,
  NA,                 2,           10,
  NA,                 3,           9,
  "Katherine Burke",  1,           4
)

treatment%>%
        fill(person)
```

###Case Study
```{r}
who
who1<- who %>%
        pivot_longer(
                cols=new_sp_m014:newrel_f65,
                names_to= "key",
                values_to= "cases",
                values_drop_na= T
        )
who1

who1 %>%
        count(key) #to explore and see if there's any patterns in values
unique(who1['key']) #to explore the unique values in key

who2 <- who1 %>% #mutating existing table with new values before separating
        mutate(key=stringr::str_replace(key,"newrel", "new_rel"))
who2

who3<- who2 %>%
        separate(key,c("new", "type", "sexage"), sep="_")
who3 %>%
        count(new)
unique(who3['new']) #see how many unique values and if column is useful

who4<- who3 %>%
        select(-iso3, -iso2, -new)
who4
```

